---
title: "Manuscript MAIN table: 57 candidate genes out of 91 TOP_perm100K_NOM"
author: "Emilio Ugalde"
date: "2023-06-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**DESCRIPTION**: \
Generate MAIN table including TWAS robust associations found as conditionally independent in the joint-conditional analysis: which is, 57 genes at 51 loci.


```{r}
# version sufix
ver = "v0.5.dat"

# Setup
twas="TWAS_testis_tgct_scca_19839uniqGENES"

## Source directory
path_machine="/run/user/1000/gvfs/sftp:host=vector.meb.ki.se"
wd="/nfs/GENETEC/TWAS/"
wd=paste0(path_machine, wd)
```

## Input
NOTE: master file created using scripts: ./Manuscript/Scripts_results/1.0.Generate_TWAS_annotated_master_table_100KpermNOM_v0.5.Rmd \

```{r}
dir=paste0(wd, "Manuscript/Tables/annot_100KpermNOM_only/")
imp="TWAS_testis_tgct_scca_19839uniqGENES_91permNOM_51LOCI.master_ANNOT.dat"
```

## Output
```{r}
dir_out=paste0(wd, "Manuscript/Tables/Formatted/")
# File names 
name_out=c("Main_Table.Perm100K_NOM_Candidate_genes")
```

```{r}
# columns to retrieve
#col_get <- c("Locus","CHR","ID","geneType","TISSUE","JOINT.Z","JOINT.P","TGCT-GWAS" )
col_get <- c("LOCUS.index","Cytoband_hg38","ID","Position_hg38","geneType","TISSUE","JOINT.Z","JOINT.P","TGCT-GWAS" )

```
```{r}
# Edit naming of variables
col_edit <- c("Locus index","Cytoband","Gene symbol","Position (hg38)","Gene type","Tissue","Joint Z-score","Joint P-value","Overlap GWAS locus")
```


## Read in data
```{r}
top <- read.table(paste0(dir, imp), header=T, sep="\t")
nrow(top)
```
## Format data

```{r}
# Re-name Tissues
top$TISSUE <- ifelse(top$PANEL=="GTExv8.ALL.Testis", "Testis", 
                     ifelse(top$PANEL=="TCGA-TGCT.TUMOR", "TGCT", "Cross-tissue"))
# Position (hg38)
top$Position_hg38 <- paste0("chr",top$CHR, ": ", paste0(top$P0_hg38, "-",top$P1_hg38))
# Position (hg19)
top$Position_hg19 <- paste0("chr",top$CHR, ": ", paste0(top$P0_hg19, "-",top$P1_hg19))
# Annotate TGCT-GWAS overlap (i.e. add "*" if gene same as prev. adjudicated-implicated)
top$`TGCT-GWAS` <- ifelse(top$GWAS_overlap=="Yes", "Y", "N")
top$`TGCT-GWAS` <- ifelse(top$GWAS_overlap=="Yes" & top$gwas_gene==T , "Y*", top$`TGCT-GWAS`)
table(top$`TGCT-GWAS`)
# Clean geneType names
top$geneType_bk <- top$geneType
top$geneType <- replace(top$geneType, top$geneType=="protein_coding", "PC")
top$geneType <- replace(top$geneType, top$geneType=="antisense", "ant")
top$geneType <- replace(top$geneType, top$geneType=="lincRNA", "lnc")
top$geneType <- replace(top$geneType, top$geneType=="processed_pseudogene", "pps")
top$geneType <- replace(top$geneType, top$geneType=="processed_transcript", "ptr")
top$geneType <- replace(top$geneType, top$geneType=="TEC", "tec")
#top$geneType <- gsub("_", " ", top$geneType)
```
```{r, eval=F}
top[, col_get]
```



```{r}
# Round decimals
top$TWAS.Z <- round(top$TWAS.Z, 4)
top$JOINT.Z <- round(top$JOINT.Z, 4)
top$JOINT.P <- formatC(top$JOINT.P, format="e", digits = 2)
```


```{r}
# Annotate if prev. implicated-adjudicated
top$Adjud_impl.status <- ifelse(top$gwas_gene==T, "*", "")
top$Adjud_impl.status <- replace(top$Adjud_impl.status, is.na(top$Adjud_impl.status)==T, "")
table(top$Adjud_impl.status)
```


# Filter data
```{r}
out <- top[top$PERM.nom=="yes", ]
nrow(out)
# Loci
length(unique(out$LOCUS))
```
- Tabulate by protein-coding type
```{r}
table(out$geneType=="protein_coding")
```

## Filter Leading genes
```{r}
table(out$JOINT_status)
```
```{r}
# out_lead <- out[out$JOINT_status=="single" | out$JOINT_status=="joint_best", ]
out_lead <- out[out$JOINT_status!="conditional" & out$JOINT_status!="correlated", ]
nrow(out_lead)
# Loci
length(unique(out_lead$LOCUS))
```
- Tabulate by GWAS-locus overlap
```{r}
table(out_lead$GWAS_overlap)
table(out_lead$GWAS_overlap_locus)
```

- Number of unique GWAS loci
```{r}
# Known loci
length(unique(out_lead[out_lead$GWAS_overlap=="Yes",]$LOCUS))
# Novel loci
length(unique(out_lead[out_lead$GWAS_overlap=="No",]$LOCUS))
```

- Tabulate by protein-coding type
```{r}
table(out_lead$geneType=="protein_coding")
```
- Tabulate by protein-coding type: known/novel loci
```{r}
table(out_lead$geneType=="protein_coding", out_lead$GWAS_overlap)
```

- Tabulate if TWAS gene is same as GWAS.implicated-adjudicated
```{r}
table(out_lead$Adjud_impl.status)
```

# Produce main table

```{r}
out_table <- out_lead[ order(out_lead$CHR, out_lead$P0_hg38), col_get]
```

```{r}
file=paste0(name_out, "_", nrow(out_lead),"genes_", ver)
#cat(append(head_out, "\n"), sep = "\t", file=paste0(dir_out, file), append = T )
cat(append(col_edit, "\n"), sep = "\t", file=paste0(dir_out, file), append = T )

write.table(out_table, file=paste0(dir_out, file), col.names = F, row.names = F, quote = F, sep="\t", append = T)

print(file)
```
```{r}
# Print table
out_table
```


```{r, eval=F}
# To create PDF table with scientific notation
library(formatdown)
top$JOINT.P <- format_power(as.numeric(top$JOINT.P), format="sci")
```
```{r, eval=F}
knitr::kable(out_lead[ order(out_lead$Locus), col_get], col.names = col_edit)
```


