---
title: "Figure 1A. Pie chart of prediction models from panels"
author: "Emilio Ugalde"
date: "April 20, 2023"
output: html_document
---

For R solution, see: https://syoxscorner.com/how-to/r/creating-pie-charts-with-r-using-ggplot2/


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2) 
library(scales)
```

**OVERVIEW**: Summary of TWAS analysis using TECAC sumstats and five gene expression prediction panels: testis tissue (GTEx v_8), TGCT tumor tissue (TCGA), and three main cross-tissue features (sparse cannonical correlation analysis, sCCA, from 22 GTEx normal tissues). Testis was considered the main target tissue. Additional prediction models were included if available in TGCT panel, or if not, from the sCCA feature with best gene-level predictive performance.


```{r, echo=T}
# Source directory
path_machine="/run/user/1000/gvfs/sftp:host=vector.meb.ki.se"
wd="/nfs/GENETEC/TWAS/"
wd=paste0(path_machine, wd)
```
```{r}
# version
ver="v0.1"
```


# Set output
```{r, echo=T}
# Output dir
dir_out=paste0(wd, "Manuscript/Figures/")
file_prefix="Figure_1A.Plot_PANEL_pred_models_"
```

# Input
See scripts: Programs/2.Summarize_TWAS_v4_output_v2.1.R \

## TWAS output files
```{r}
# Formatted gene names and hg19 coordinates
dir=paste0(wd, "Output/imputed_TECAC/summary_perm100K/")
all <- read.table(file=, paste0(dir, 'TWAS_testis_tgct_scca_19839uniqGENES.dat'), header=T, sep="\t")
```

## Tabulate prediction models

- Mutually exclusive models
```{r}
# Re-name panels
all$PANEL_x <- replace(all$PANEL, all$PANEL=="sCCA1" | all$PANEL=="sCCA2" | all$PANEL=="sCCA3", "Cross-tissue")
all$PANEL_x[all$PANEL_x=="GTExv8.ALL.Testis" ] <- "Testis"
all$PANEL_x[all$PANEL_x=="TCGA-TGCT.TUMOR" ] <- "TGCT"
table(all$PANEL_x)
```

```{r}
tab_models <- merge(as.data.frame(table(all[all$PANEL_x=="Testis", ]$MODEL)), as.data.frame(table(all[all$PANEL_x=="TGCT", ]$MODEL)), by="Var1", all=T)
tab_models <- merge(tab_models, as.data.frame(table(all[all$PANEL_x=="Cross-tissue", ]$MODEL)), by="Var1", all=T)
names(tab_models) <- c("Model", "Testis", "TGCT", "Cross-tissue" )
tab_models[is.na(tab_models)] <- "-"
tab_models
```

```{r}
knitr::kable(tab_models, align='c')
```
```{r}
knitr::kable(tab_models, format = "latex", align='c')
```

```{r}
write.table(tab_models, file=paste0(dir_out, "Fig1A_MODELS_Tabulated.dat"), row.names=F, col.names=T, quote=F, sep="\t")
```

```{r, eval=F}
library(knitr)
#library(dplyr)
library(kableExtra)
kable(tab_models, "html") %>% save_kable(file=paste0(dir_out, "VenDiagram_MODELS_Tabulated.pdf"))
pdf(file=paste0(dir_out, "VenDiagram_MODELS_Tabulated.pdf"), width = 4, height = 3)
knitr::kable(tab_models)
dev.off()
```


## Pie chart
```{r}
t_order=c("Testis", "Cross-tissue", "TGCT")
tb <- as.data.frame(table(all$PANEL_x) )
tb$Var1 <- factor(as.character(tb$Var1), levels = t_order)
names(tb) <- c("Tissue", "n")
tb$prop <- paste0(round(prop.table(tb[,2])*100, 1), "%")
#rownames(tb) <- tb$Tissue
#tb <- tb[t_order, ]
```

```{r, eval=F}
ggplot(tb, aes(x="", y=n, fill=Tissue)) +
  geom_bar(width = 1, stat = "identity", color="white") +
  coord_polar(theta="y", start=0) +
  theme_void() +
  geom_text(aes(
    y = c(3617.5, 13331.0, 19633.0), # n / 2 + c(0, cumsum(n)[-length(n)]
    label=prop), size=3)
```

```{r}
# Add label (%) positions
library(dplyr)
num.data <- tb %>% 
  arrange(desc(Tissue)) %>%
  mutate(lab.ypos = cumsum(n) - 0.5*n)
num.data
```

# Plot
```{r}
ggplot(num.data, aes(x="", y=n, fill=Tissue)) +
  geom_bar(width = 1, stat = "identity", color="white") +
  coord_polar(theta="y", start=0) +
    geom_text(aes(
    y = lab.ypos, # n / 2 + c(0, cumsum(n)[-length(n)]
    label=prop), size=5, color="black") +
  theme_void() 
```

# Donut plot
```{r}
ggplot(num.data, aes(x=2, y=n, fill=Tissue)) +
  geom_bar(width = 1, stat = "identity", color="white") +
  coord_polar(theta="y", start=0) +
    geom_text(aes(
    y = lab.ypos, # n / 2 + c(0, cumsum(n)[-length(n)]
    label=prop), size=5, color="black") +
  theme_void() + theme(legend.title = element_text(face="bold"), legend.text = element_text(size=13) ) +xlim(0.5,2.5) 
```

## Remove percentage on TGCT
```{r}
num.data_2 <- num.data
num.data_2[num.data_2$Tissue=="TGCT",3] <- " "
```

```{r}
pd1 <- ggplot(num.data_2, aes(x=2, y=n, fill=Tissue)) +
  geom_bar(width = 1, stat = "identity", color="white") +
  coord_polar(theta="y", start=0) +
    geom_text(aes(
    y = lab.ypos, # n / 2 + c(0, cumsum(n)[-length(n)]
    label=prop), size=6, color="black") +
  theme_void() + theme(legend.title = element_blank(), legend.text = element_text(size=20) ) +xlim(0.5,2.5)  
pd1
```
```{r}
# legend at the bottom
pd1_b <- ggplot(num.data_2, aes(x=2, y=n, fill=Tissue)) +
  geom_bar(width = 1, stat = "identity", color="white") +
  coord_polar(theta="y", start=0) +
    geom_text(aes(
    y = lab.ypos, # n / 2 + c(0, cumsum(n)[-length(n)]
    label=prop), size=6, color="black") +
  theme_void() + theme(legend.title = element_blank(), legend.text = element_text(size=16), legend.position = "bottom" ) +xlim(0.5,2.5)  
pd1_b
```
```{r}
# legend at the top
pd1_u <- ggplot(num.data_2, aes(x=2, y=n, fill=Tissue)) +
  geom_bar(width = 1, stat = "identity", color="white") +
  coord_polar(theta="y", start=0) +
    geom_text(aes(
    y = lab.ypos, # n / 2 + c(0, cumsum(n)[-length(n)]
    label=prop), size=5, color="black") +
  theme_void() + theme(legend.title = element_blank(), legend.text = element_text(size=14), legend.position = "top" ) +xlim(0.5,2.5)  
pd1_u
```

# Save plots
```{r}
file_out=paste0(file_prefix, ver, ".pdf")
file_out
pdf(file=paste0(dir_out, file_out), width=6, height=5)
pd1
dev.off()
```

```{r}
file_out=paste0(file_prefix, "BottomLegend_" , ver, ".pdf")
file_out
pdf(file=paste0(dir_out, file_out), width=4, height=5)
pd1_b
dev.off()
```
```{r}
file_out=paste0(file_prefix, "TopLegend_" , ver, ".pdf")
file_out
pdf(file=paste0(dir_out, file_out), width=4, height=5 )
pd1_u
dev.off()
```

```{r, eval=F}
# PNG format
file_out=paste0(file_prefix , ver, ".png")
file_out
png(file=paste0(dir_out, file_out), width=5, height=5, units = "in", res=300 )
pd1
dev.off()
```

```{r, eval=F}
file_out=paste0(file_prefix, "TopLegend_" , ver, ".png")
file_out
png(file=paste0(dir_out, file_out), width=6, height=7, units = "in", res=300 )
pd1_u
dev.off()
```

