---
title: "Model predictive performance"
author: "Emilio Ugalde"
date: "2022-10-24"
output: html_document
---

Description (v0.3): xlimits set between 0 to 1 and using same type of bar-bin size (breaks=seq(0,1,0.04)). \

```{r setup, include=FALSE}
library(plyr)
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
# Source directory
path_machine="/run/user/1000/gvfs/sftp:host=vector.meb.ki.se"
wd="/nfs/GENETEC/TWAS/"
wd=paste0(path_machine, wd)
```

## Set output
```{r}
# version
ver="_v0.3"
# Output dir
dir_out=paste0(wd, "Manuscript/Figures/")
file_out=c("Suppl.Figure_1A.Histograms_models_heritability", "Suppl.Figure_1B.Histograms_models_accuracy" )
```

## Input
```{r}
dir="/nfs/GENETEC/TWAS/Output/imputed_TECAC/summary_perm100K/"
all <- read.table(paste0(path_machine, dir, 'TWAS_testis_tgct_scca_19839uniqGENES.dat'), header=T, sep="\t" )
nrow(all)
```
```{r}
# Re-name panels
all$PANEL_x <- replace(all$PANEL, all$PANEL=="sCCA1" | all$PANEL=="sCCA2" | all$PANEL=="sCCA3", "Cross-tissue")
all$PANEL_x[all$PANEL_x=="GTExv8.ALL.Testis" ] <- "Testis_normal"
all$PANEL_x[all$PANEL_x=="TCGA-TGCT.TUMOR" ] <- "TGCT_tumor"
```

## R2 P-value distribution
```{r}
# all models
table(cut(all$MODELCV.PV, c(0,0.001, 0.01, 0.05, 0.1,0.5,1) ) )
```


## Heritability distribution

- All panesl
```{r}
summary(all$HSQ)
```

- Per panel
```{r}
d_hsq <- ddply(all, "PANEL_x", summarise, grp.mean=mean(HSQ))
d_hsq
```


## R2 distribution

- All panels
```{r}
# All models
summary(all[all$MODELCV.PV<0.05,]$MODELCV.R2)
```
- Per panel
```{r}
d_r2 <- ddply(all, "PANEL_x", summarise, grp.mean=mean(MODELCV.R2))
d_r2
```

## Prediction accuracy (R2 standardized by SNP heritability)
```{r, echo=F}
all$CV_acu <- all$MODELCV.R2 / all$HSQ
summary(all$CV_acu)
```
- Per panel
```{r}
d_acu <- ddply(all, "PANEL_x", summarise, grp.mean=mean(CV_acu))
d_acu
```


## Prediction accuracy (R2 standardized by SNP heritability): upper-bounded
```{r}
# trick, set to 1 if>1
all$CV_acu_1 <- ifelse(all$CV_acu>1, 1, all$CV_acu)
summary(all$CV_acu_1)
```
- Per panel
```{r}
d_acu1 <- ddply(all, "PANEL_x", summarise, grp.mean=mean(CV_acu_1))
d_acu1
```

\newpage
# Plots

## Histogram cis-H2 heritability
```{r}
par(mfrow=c(2,2))
# all panels
hist(all$HSQ, main='All', xlab = NULL, ylab = "No. of models", breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all$HSQ), col = "darkred")
# Testis panel
hist(all[all$PANEL_x=='Testis_normal',]$HSQ, main='Normal testis', xlab = NULL, ylab = NULL, breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='Testis_normal',]$HSQ), col = "darkred")
# Cross-tissue panels
hist(all[all$PANEL_x=='Cross-tissue',]$HSQ, main='Cross-tissue', xlab = "SNP heritablity ", ylab = "No. of models", breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='Cross-tissue',]$HSQ), col = "darkred")
# TGCT tumor panel
hist(all[all$PANEL_x=='TGCT_tumor',]$HSQ, main='TGCT', xlab = "SNP heritablity", ylab = NULL, breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='TGCT_tumor',]$HSQ), col = "darkred")
```

- Save
```{r, echo=T}
png(file=paste0(dir_out, file_out[1], ver, ".png"), width=9, height=6, units="in", res=600 )

par(mfrow=c(2,2))
# all panels
hist(all$HSQ, main='All', xlab = NULL, ylab = "No. of models", breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all$HSQ), col = "darkred")
# Testis panel
hist(all[all$PANEL_x=='Testis_normal',]$HSQ, main='Normal testis', xlab = NULL, ylab = NULL, breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='Testis_normal',]$HSQ), col = "darkred")
# Cross-tissue panels
hist(all[all$PANEL_x=='Cross-tissue',]$HSQ, main='Cross-tissue', xlab = "SNP heritablity ", ylab = "No. of models", breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='Cross-tissue',]$HSQ), col = "darkred")
# TGCT tumor panel
hist(all[all$PANEL_x=='TGCT_tumor',]$HSQ, main='TGCT', xlab = "SNP heritablity", ylab = NULL, breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='TGCT_tumor',]$HSQ), col = "darkred")

dev.off()
```



## Histogram variance explained (R2)
```{r, echo=T}
par(mfrow=c(2,2))
# all panels
hist(all$MODELCV.R2, main='All', xlab = NULL, ylab = "No. of models", breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all$MODELCV.R2), col = "darkred")
# Testis panel
hist(all[all$PANEL_x=='Testis_normal',]$MODELCV.R2, main='Testis', xlab = NULL, ylab = NULL, breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='Testis_normal',]$MODELCV.R2), col = "darkred")
# Cross-tissue panels
hist(all[all$PANEL_x=='Cross-tissue',]$MODELCV.R2, main='Cross-tissue', xlab = "Variance explained R2", ylab = "No. of models", breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='Cross-tissue',]$MODELCV.R2), col = "darkred")
# TGCT tumor panel
hist(all[all$PANEL_x=='TGCT_tumor',]$MODELCV.R2, main='TGCT_tumor', xlab = "Variance explained R2", ylab = NULL, breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='TGCT_tumor',]$MODELCV.R2), col = "darkred")
```



## Prediction accuracy

```{r, echo=T}
par(mfrow=c(2,2))
# all panels
hist(all$CV_acu, main='All', xlab = NULL, ylab = "No. of models", breaks = "Sturges")
abline(v = mean(all$CV_acu), col = "darkred")
# Testis panel
hist(all[all$PANEL_x=='Testis_normal',]$CV_acu, main='Normal testis', xlab = NULL, ylab = NULL, breaks = "Sturges")
abline(v = mean(all[all$PANEL_x=='Testis_normal',]$CV_acu), col = "darkred")
# Cross-tissue panels
hist(all[all$PANEL_x=='Cross-tissue',]$CV_acu, main='Cross-tissue', xlab = "Accuracy R2/h2g", ylab = "No. of models", breaks = "Sturges")
abline(v = mean(all[all$PANEL_x=='Cross-tissue',]$CV_acu), col = "darkred")
# TGCT tumor panel
hist(all[all$PANEL_x=='TGCT_tumor',]$CV_acu, main='TGCT_tumor', xlab = "Accuracy R2/h2g", ylab = NULL, breaks = "Sturges")
abline(v = mean(all[all$PANEL_x=='TGCT_tumor',]$CV_acu), col = "darkred")
```

## Prediction accuracy (upper bounded by heritability)
```{r}
par(mfrow=c(2,2))
# all panels
hist(all$CV_acu_1, main='All', xlab = NULL, ylab = "No. of models", breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all$CV_acu_1), col = "darkred")
# Testis panel
hist(all[all$PANEL_x=='Testis_normal',]$CV_acu_1, main='Normal testis', xlab = NULL, ylab = NULL, breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='Testis_normal',]$CV_acu_1), col = "darkred")
# Cross-tissue panels
hist(all[all$PANEL_x=='Cross-tissue',]$CV_acu_1, main='Cross-tissue', xlab = "Accuracy R2/h2g", ylab = "No. of models", breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='Cross-tissue',]$CV_acu_1), col = "darkred")
# TGCT tumor panel
hist(all[all$PANEL_x=='TGCT_tumor',]$CV_acu_1, main='TGCT', xlab = "Accuracy R2/h2g", ylab = NULL, breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='TGCT_tumor',]$CV_acu_1), col = "darkred")
```

- Save
```{r, echo=T}
png(file=paste0(dir_out, file_out[2], ver, ".png"), width=9, height=6, units="in", res=600 )

par(mfrow=c(2,2))
# all panels
hist(all$CV_acu_1, main='All', xlab = NULL, ylab = "No. of models", breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all$CV_acu_1), col = "darkred")
# Testis panel
hist(all[all$PANEL_x=='Testis_normal',]$CV_acu_1, main='Normal testis', xlab = NULL, ylab = NULL, breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='Testis_normal',]$CV_acu_1), col = "darkred")
# Cross-tissue panels
hist(all[all$PANEL_x=='Cross-tissue',]$CV_acu_1, main='Cross-tissue', xlab = "Accuracy R2/h2g", ylab = "No. of models", breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='Cross-tissue',]$CV_acu_1), col = "darkred")
# TGCT tumor panel
hist(all[all$PANEL_x=='TGCT_tumor',]$CV_acu_1, main='TGCT', xlab = "Accuracy R2/h2g", ylab = NULL, breaks = seq(0,1,0.04), xlim=c(0,1))
abline(v = mean(all[all$PANEL_x=='TGCT_tumor',]$CV_acu_1), col = "darkred")

dev.off()
```
















