---
title: "Germ cell neoplasia in situ (GCNIS) dataset"
author: "Emilio Ugalde"
date: "2023-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
library(tidyverse)
```

```{r}
dir="~/Documents/postdoc_FW/projects/TWAS/CNIS"
```

## Output
- Gene-level aggreagated dataset on relevant cell-types
```{r}
file_out="/DerivedData/GCNIS_aggregated_20230314.txt"
```

## INPUT
- GCNIS Sone dataset
```{r}
read.delim(paste0(dir, "/files_Kristian/OutData_GL.txt")) ->allData
#read.delim(paste0(dir, "/files_Kristian/TECAC_TWAS_matrix.txt",header = FALSE))->MA1
```

- Gene ids
```{r}
hgnc <- data.table::fread(paste0(dir, "/annot/HGNC_genenames.txt"))
```
```{r}
hgnc$ID <- hgnc$`Approved symbol`
hgnc$entrez <- hgnc$`NCBI Gene ID`
```

- TWAS results
```{r}
dir_twas <- "/home/emiuga/Documents/postdoc_FW/projects/TWAS/Manuscript/dropbox_v2/"
file_twas <- "TWAS_testis_tgct_scca_19839uniqGENES_165FDR01_91permNOM_51LOCI.master_ANNOT.dat"
#file_twas_full <- "TWAS_testis_tgct_scca_19839uniqGENES.dat"
twas <- read.table(paste0(dir_twas, file_twas), header=T, sep="\t" )
#full <- read.table(paste0(dir_twas, file_twas_full), header=T, sep="\t" )
```

## Get list of TWAS genes
```{r}
# Top genes (165 FDR<0.01)
tlist_top <- twas$ID
```

- Directionality (excludding 'Cross-tissue' panel)
```{r}
summary(as.factor(sign(twas[twas$PANEL_x!="Cross-tissue", ]$TWAS.Z) ))
```
- Up-regulated
```{r}
tlist_top_UP <- twas[twas$PANEL_x!="Cross-tissue" & sign(twas$TWAS.Z)>0, ]$ID
str(tlist_top_UP)
```
- Dow-regulated
```{r}
tlist_top_DN <- twas[twas$PANEL_x!="Cross-tissue" & sign(twas$TWAS.Z)<0, ]$ID
str(tlist_top_DN)
```


## Format data
```{r}
# NOTE: using updated ID, match with TWAS went from 105 to 117 genes
# Merge by ENTREZ id
allData_m <- merge(allData[-c(1:3),-c(1,2,4)], hgnc[is.na(hgnc$entrez)==F, c("ID", "entrez")], by="entrez") 
table(allData_m$symbol==allData_m$ID)
table(twas$ID%in%allData_m$symbol)
table(twas$ID%in%allData_m$ID)
```


- Retrieve expression data for genes with valid ENTREZ id
```{r}
allData_g <- merge(allData[-c(1:3),-c(1,2,4,5)], hgnc[is.na(hgnc$entrez)==F, c("ID", "entrez")], by="entrez") 
cell_names <- as.character(allData[1,-c(1,2,4,5)])
cell_names[1] <- "ID"
col_lim <- ncol(allData_g)-1
allData_g <- allData_g[, c(ncol(allData_g),2:col_lim) ]
names(allData_g) <- cell_names
dim(allData_g)
```
## Heatmap

```{r, eval=F}
### Plot From Kristian
### NOTE: seems it has bugs, did not succeed to reproduce!
MA1[,-c(1:5)]->MA
MA[,-c(2:29)]->MA
MA[-c(1,3,4),]->MA

MA->MA2

colnames(MA2)<-MA2[1,]
MA2[-1,]->MA2
#colnames(MA)<-c("Symbol", "ESC1","ESC2", "ESC3","FetOv1","FetOv2","FetOv3","FetGon1","FetGon2", "FetGon3","CIS1","CIS2","CIS3","CIS_Total1","CIS_Total2","CIS_Total3","NT1","NT2","NT3", "Spgenesis1", "Spgenesis2")
as.numeric(unlist(MA2[,-1]))->MA2[,-1]
#as_tibble(MA)
rownames(MA2)<-MA2$Cy3

data.matrix(MA2[,-1])->MAmatrix
heatmap(MAmatrix, Colv = NA, scale="none", cexRow=1, labRow=paste(MA[,1]), col= colorRampPalette(brewer.pal(8, "Blues"))(25))
```

```{r}
gl15 <- c("EXOG", "GINM1", "HEATR3", "CRYL1", "USP36", "MCM3AP", "TMEM52B", "RFWD3", "CAPNS1", "UCK2", "OLR1", "ZFPM1", "ZBTB9", "LZTR1")

hm1 <- allData_g[allData_g$ID%in%gl15, ]
hm1[,-1] <- as.numeric(unlist(hm1[,-1]))
hm1_m <- data.matrix(hm1[, -1]) 
#hm1_m <- as.matrix(hm1[, -1]) 
heatmap(hm1_m, Colv = NA, scale="column", cexRow=1.0, labRow=paste(hm1[, 1]), col= colorRampPalette(brewer.pal(8, "Blues"))(25))
```

## Generate FULL dataset with relevant cell-types: enbryonic stem cells (ESC), GCNIS, and Spermatogenesis

```{r}
cell_g <- allData_g[,c(1:4,11:21)]
cell_g[,-1] <- as.numeric(unlist(cell_g[,-1]))
#cell_gM <- data.matrix(cell_g[, -1]) 

# Collapsed data set: mean expression across cell types
cell_g_mean <- data.frame(ID=cell_g[,1], ESC=rowMeans(cell_g[,2:4]),GCNIS=rowMeans(cell_g[,5:10]),Spermatogenesis=rowMeans(cell_g[,11:15]))
```


```{r}
# Aggregate by MEAN gene-level expression
cell_g_agg <- aggregate(x = cell_g_mean[,2:4], by = list(cell_g_mean$ID), FUN = mean)
names(cell_g_agg)[1] <- "ID"
dim(cell_g_agg)
```

# Save aggregated dataset
```{r}
file_out
write.table(cell_g_agg, file=paste0(dir, file_out), col.names=T, row.names=F, quote=F, sep="\t" )
```

# Explore visualization

## Dot pot: selected 15 top genes
```{r}
glist15 <- c("EXOG", "GINM1", "HEATR3", "CRYL1", "USP36", "MCM3AP", "TMEM52B", "RFWD3", "CAPNS1", "UCK2", "OLR1", "ZFPM1", "ZBTB9", "LZTR1")
str(glist15)
```
```{r}
# Subset eset
e_top15 <- cell_g_agg[cell_g_agg$ID%in%glist15, ]
dim(e_top15)
```
```{r}
## EUM: scale expression of genes per cell-type level
e_top15_s <- e_top15
e_top15_s[,-1] <- apply(e_top15_s[,-1 ], 2, function(x) {scale(x, center=F)})
```

```{r}
e_top15_long <- gather(e_top15_s, condition, ave_exp, ESC:Spermatogenesis,factor_key=TRUE)
```
```{r}
e_top15_long <- e_top15_long %>%
  mutate(ID=ID %>% fct_relevel("EXOG", "GINM1", "HEATR3", "CRYL1", "USP36", "MCM3AP", "TMEM52B", "RFWD3", "CAPNS1", "UCK2", "OLR1", "ZFPM1", "ZBTB9", "LZTR1"))
```

```{r}
ggplot(e_top15_long, aes(x=ID, y=condition, size=ave_exp, color=ave_exp)) + geom_point(alpha = 0.8) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +  labs(x="",y="")+scale_color_gradient2(low = "wheat1", mid = "chocolate1", high = "firebrick4", midpoint=round(mean(e_top15_long$ave_exp), 1), space = "Lab")
```

```{r}
ggplot(e_top15_long, aes(x=ID, y=condition, size=ave_exp, color=ave_exp)) + geom_point(alpha = 0.8) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +  labs(x="",y="")+scale_color_gradient2(low = "lightskyblue1", mid = "turquoise1", high = "royalblue4", midpoint=round(mean(e_top15_long$ave_exp), 1), space = "Lab")
```

