---
title: "Expression enrichment analysis (TWAS joint independent genes): Germ cell neoplasia in situ (GCNIS) dataset"
author: "Emilio Ugalde"
date: "April 19, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
library(tidyverse)
```

```{r}
dir="~/Documents/postdoc_FW/projects/TWAS/CNIS"
ver="_v0.3"
```


## Output file
```{r}
# File path and prefix
file_out=paste0("../Output/PermuPV_GCNIS", ver )
```

## INPUT
- Gene-level aggregated data set on relevant conditions
see: Scripts/Process_GCNIS_data_v0.1.Rmd
```{r}
# file prefix
file="/DerivedData/GCNIS_aggregated_20230314.txt"
```

```{r}
read.table(paste0(dir, file), header = T) -> allData_agg
```


- TWAS results
```{r}
dir_twas <- "/home/emiuga/Documents/postdoc_FW/projects/TWAS/Manuscript/dropbox_v2/"
file_twas <- "TWAS_testis_tgct_scca_19839uniqGENES_165FDR01_91permNOM_51LOCI.master_ANNOT.dat"
file_twas_full <- "TWAS_testis_tgct_scca_19839uniqGENES.dat"
twas <- read.table(paste0(dir_twas, file_twas), header=T, sep="\t" )
full <- read.table(paste0(dir_twas, file_twas_full), header=T, sep="\t" )
```

## Distribution
```{r}
apply(allData_agg[,-1], 2, summary)
```

## Format data (log10 scale)
```{r}
allData_agg_S <- allData_agg
allData_agg_S[,-1] <- log10(allData_agg[,-1]+1)
```
```{r}
apply(allData_agg_S[,-1], 2, summary)
```

# Analysis (top TWAS)

## Get list of TWAS genes
```{r}
# Jointly independent (57 genes)
tlist_top <- twas[which(twas$JOINT), ]$ID
str(tlist_top)
```

## TWAS subset (JOINT independent genes)
```{r}
e_twas_top <- allData_agg_S[allData_agg_S$ID%in%tlist_top,] 
nrow(e_twas_top)
```

## Calculate observed values for the TWAS gene set
```{r}
# Scaled Median AVERAGE expression
med_top_M <- apply(e_twas_top[,-1], 2, median ) 
med_top_M
```

## Background set (genes included in TWAS)
```{r}
glist_bk <- allData_agg_S$ID[allData_agg_S$ID%in%full$ID]
str(glist_bk)
```

- Subset data set
```{r}
e_twas_bk <- allData_agg_S[allData_agg_S$ID%in%glist_bk,]
dim(e_twas_bk)
```

- Median expression
```{r}
median_e_bk <- apply(e_twas_bk[,-1], 2, median )
median_e_bk
```

- Median expression in FULL dataset
```{r}
apply(allData_agg_S[,-1], 2, median ) 
```

\newpage
# Permutation tests
- Parameters
```{r}
seed=15684
set.seed(seed)
perm=10000
```

## Estimate MEDIAN average expression, SCALED (at a given cell-type) for a random set of genes

```{r}
# Data set SIZE
s_size=nrow(e_twas_top)
n_cat <- 3
nrow(e_twas_top)
```


```{r, include=T, echo=T}
X_twas <- e_twas_bk[, -1]
# Run permutations
P_ave_sc_median <- matrix(rep(NA, n_cat*perm), ncol=n_cat)
for(i in 1:perm) {
  s <- X_twas[ sample(nrow(X_twas), size=s_size),  ]
  P_ave_sc_median[i,] <- apply(s, 2, median )
}
```
- Distribution in random sets
```{r}
apply(P_ave_sc_median, 2, summary)
```

- Calculate perm-Pval: no. of times median average expression in random sets is equal or equal than the observed. (per condition)
```{r}
# Compute empirical p-value for each cell-type
med_ave_sc_pval <- rep(NA, n_cat)
for(i in 1:n_cat) {
med_ave_sc_pval[i] <- ( sum(P_ave_sc_median[ ,i]>=med_top_M[i]) + 1  ) / ( perm+1 )
}
```
```{r}
names(med_ave_sc_pval) <- names(allData_agg[, -1])
med_ave_sc_pval
```
## Compile results
```{r}
# Set order
res_order <- c("ESC", "Spermatogenesis", "GCNIS")
```

```{r}
out <- data.frame(Condition=names(med_ave_sc_pval[res_order]),
                  Background=c(nrow(e_twas_bk), " ", " "), Leading=c(nrow(e_twas_top), " ", " "), # Overlapping set (no. of genes)
                  Background=median_e_bk[res_order], Leading=med_top_M[res_order], # Median average expression
                  PERM.PV=med_ave_sc_pval[res_order])
out
```

## Save 
```{r}
file=paste0(file_out, ".dat")
write.table(out, file=file, col.names=T, row.names=F, quote=F, sep="\t" )
file
```
