---
title: "Expression enrichment analysis (TWAS joint independent genes): Tumor expression dataset (TCGA-TGCT)"
author: "Emilio Ugalde"
date: "April 19, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#library(RColorBrewer)
#library(tidyverse)
```

```{r}
dir="~/Documents/postdoc_FW/projects/TWAS/TCGA_tgct/"
ver="_v0.3"
```

## Output file
```{r}
# File path and prefix
file_out=paste0("../Data/PermuPV_TGCT-TCGA", ver )
```

## INPUT
See: Scripts/Process_TCGAtgct_data_v0.2.Rmd \
```{r}
file="Data/norm_mat_150.aggregated_20230417.txt"
```

```{r}
# library(data.table)
# fread(paste0(dir, file)) -> eset
read.delim(paste0(dir, file)) -> eset
# Remove ensembl.id column
eset <- eset[,-1]
```

- TWAS results
```{r}
dir_twas <- "/home/emiuga/Documents/postdoc_FW/projects/TWAS/Manuscript/Draft/Results_20230203/"
file_twas <- "TWAS_testis_tgct_scca_19839uniqGENES_91permNOM_51LOCI.master_ANNOT.dat"
file_twas_full <- "TWAS_testis_tgct_scca_19839uniqGENES.dat"
twas <- read.table(paste0(dir_twas, file_twas), header=T, sep="\t" )
full <- read.table(paste0(dir_twas, file_twas_full), header=T, sep="\t" )
```

## Overlap with TWAS
```{r}
table(twas$ID%in%eset$ID)
table(full$ID%in%eset$ID)
```
## Distribution
```{r}
summary(eset[,-1])
```

## Format data

- Filter out non-informative genes
```{r}
# Remove entries with no gene ID
eset_f <- eset[is.na(eset$ID)==F, ]
nrow(eset_f)
# No. of unique IDs
length(unique(eset_f$ID))
```
- Remove duplicated IDs (all entries)
```{r}
dup <- unique(eset[which(duplicated(eset$ID)), ]$ID)
str(dup)
```
```{r}
eset_f <- eset_f[!eset_f$ID%in%dup, ]
nrow(eset_f)
```

## Subet on ALL genes included in TWAS analysis
```{r}
eset_T <- eset_f[eset_f$ID%in%full$ID, ]
nrow(eset_T)
```
- Distribution
```{r}
apply(eset_T[,-1], 2, summary)
```

## TWAS subset (JOINT independent genes)
```{r}
# Jointly independent (57 genes)
tlist_top <- twas[which(twas$JOINT), ]$ID
str(tlist_top)
```
```{r}
# Scaled data
e_twas_top <- eset_T[eset_T$ID%in%tlist_top,] 
nrow(e_twas_top)
```
- Genes not found
```{r}
twas[which(twas$JOINT), ][!twas[which(twas$JOINT), ]$ID%in%eset$ID, c("ID", "PANEL", "geneType", "ENSEMBL")]
```


## Calculate observed values for the TWAS gene set
```{r}
# Scaled Median AVERAGE expression
med_top_M <- apply(e_twas_top[,-1], 2, median) #median(e_twas_top$norm_exp)
med_top_M
```

## Background set (genes included in TWAS)

- Median expression
```{r}
median_e_bk <-  apply(eset_T[,-1], 2, median) 
median_e_bk
```
- Median expression in FULL dataset
```{r}
#median(as.numeric(scale(eset_S$norm_exp, center=F) ))
apply(eset_f[,-1], 2, median)
```

\newpage
# Permutation tests
- Parameters
```{r}
seed=15684
set.seed(seed)
perm=10000
```

```{r}
e_twas_bk <- eset_T
nrow(e_twas_bk)
```

## Estimate MEDIAN average expression by random sampling
```{r}
# Data set SIZE
s_size=nrow(e_twas_top)
n_cat <- 2
s_size
```

```{r}
# Run permutations
P_ave_median <- matrix(rep(NA, n_cat*perm), ncol=n_cat)
X_twas <- e_twas_bk[, -1]
for(i in 1:perm) {
  s <- X_twas[ sample(nrow(X_twas), size=s_size),  ]
  P_ave_median[i,] <- apply(s, 2, median)
}
apply(P_ave_median, 2, summary)
```

- Calculate perm-Pval: no. of times median average expression in random sets is equal or equal than the observed. (per cell-type)
```{r}
# Compute empirical p-value for each cell-type
med_ave_pval <- rep(NA, n_cat)
for(i in 1:n_cat) {
med_ave_pval[i] <- ( sum(P_ave_median[ ,i]>=med_top_M[i]) + 1  ) / ( perm+1 )
}
```
```{r}
names(med_ave_pval) <- names(e_twas_bk[, -1])
med_ave_pval
```

## Compile results
```{r}
out <- data.frame(Condition=c("Seminoma", "Non-seminoma"),
                  Background=c(nrow(eset_T), " "), Leading=c(nrow(e_twas_top), " "), # Overlapping set (no. of genes)
                  Background=median_e_bk, Leading=med_top_M, # Median average expression
                  PERM.PV=med_ave_pval)
out
```

## Save 
```{r}
file=paste0(file_out, ".dat")
write.table(out, file=file, col.names=T, row.names=F, quote=F, sep="\t" )
file
```



